### 2014-11-05 doc/thesis/evaluation.tex
Adding evaluation data for Lua Standard Libraries 5.3 (beta) and TLC

\begin{table}[!ht]
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|}
\hline
\textbf{Case study} & \textbf{easy} & \textbf{poly} & \textbf{hard} & \textbf{Total} & \textbf{\%} \\
\hline
Lua Standard Libraries & 94 & 5 & 34 & 133 & 71\% \\
\hline
MD5 & 13 & 0 & 0 & 13 & 100\% \\
\hline
LuaSocket & 109 & 1 & 13 & 123 & 89\% \\
\hline
HTTP Digest & 0 & 0 & 1 & 1 & 0\% \\
\hline
Typical & 1 & 0 & 0 & 1 & 100\% \\
\hline
Modulo 11 & 7 & 0 & 2 & 9 & 78\% \\
\hline
\end{tabular}
\end{center}
\caption{Evaluation results for each case study}
\label{tab:evalbycase}
\end{table}

@@

\begin{table}[!ht]
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|}
\hline
\textbf{Case study} & \textbf{Module} & \textbf{easy} & \textbf{poly} & \textbf{hard} & \textbf{Total} & \textbf{\%} \\
\hline
\multirow{9}{*}{Lua Standard Libraries}
& base & 9 & 1 & 16 & 26 & 35\% \\
\cline{2-7}
& coroutine & 0 & 0 & 6 & 6 & 0\% \\
\cline{2-7}
& package & 5 & 0 & 3 & 8 & 62\% \\
\cline{2-7}
& string & 14 & 0 & 0 & 14 & 100\% \\
\cline{2-7}
& table & 1 & 4 & 1 & 6 & 17\% \\
\cline{2-7}
& math & 28 & 0 & 1 & 29 & 97\% \\
\cline{2-7}
& bit32 & 12 & 0 & 0 & 12 & 100\% \\
\cline{2-7}
& io & 15 & 0 & 6 & 21 & 71\% \\
\cline{2-7}
& os & 10 & 0 & 1 & 11 & 91\% \\
\hline
\end{tabular}
\end{center}
\caption{Evaluation results for Lua Standard Libraries}
\label{tab:evallsl}
\end{table}

@@

@@

\begin{itemize}
\item \citep{tidallock}
\item \citep{bonnaire-sergeant2012typed-clojure}
\item \citep{vitousek2014deg}
\item \citep{allende2013gts}
\item \citep{tobin-hochstadt2008ts} 
\item \citep{dart}
\item \citep{typescript}
\item \citep{politz2012semantics}
\end{itemize}

>>>>

\begin{table}[!ht]
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|}
\hline
\textbf{Case study} & \textbf{easy} & \textbf{poly} & \textbf{hard} & \textbf{Total} & \textbf{\%} \\
\hline
Lua Standard Libraries 5.2 & 94 & 5 & 34 & 133 & 71\% \\
\hline
Lua Standard Libraries 5.3 & 87 & 6 & 36 & 129 & 67\% \\
\hline
MD5 & 13 & 0 & 0 & 13 & 100\% \\
\hline
LuaSocket & 109 & 1 & 13 & 123 & 89\% \\
\hline
HTTP Digest & 0 & 0 & 1 & 1 & 0\% \\
\hline
Typical & 1 & 0 & 0 & 1 & 100\% \\
\hline
Modulo 11 & 7 & 0 & 2 & 9 & 78\% \\
\hline
Typed Lua Compiler & 144 & 0 & 10 & 154 & 94\% \\
\hline
\end{tabular}
\end{center}
\caption{Evaluation results for each case study}
\label{tab:evalbycase}
\end{table}

@@

\begin{table}[!ht]
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|}
\hline
\textbf{Case study} & \textbf{Module} & \textbf{easy} & \textbf{poly} & \textbf{hard} & \textbf{Total} & \textbf{\%} \\
\hline
\multirow{9}{*}{Lua Standard Libraries}
& base & 9 & 1 & 16 & 26 & 35\% \\
\cline{2-7}
& coroutine & 0 & 0 & 6 & 6 & 0\% \\
\cline{2-7}
& package & 5 & 0 & 3 & 8 & 62\% \\
\cline{2-7}
& string & 14 & 0 & 0 & 14 & 100\% \\
\cline{2-7}
& table & 1 & 4 & 1 & 6 & 17\% \\
\cline{2-7}
& math & 28 & 0 & 1 & 29 & 97\% \\
\cline{2-7}
& bit32 & 12 & 0 & 0 & 12 & 100\% \\
\cline{2-7}
& io & 15 & 0 & 6 & 21 & 71\% \\
\cline{2-7}
& os & 10 & 0 & 1 & 11 & 91\% \\
\hline
\end{tabular}
\end{center}
\caption{Evaluation results for Lua Standard Libraries 5.2}
\label{tab:evallsl}
\end{table}

\begin{table}[!ht]
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|}
\hline
\textbf{Case study} & \textbf{Module} & \textbf{easy} & \textbf{poly} & \textbf{hard} & \textbf{Total} & \textbf{\%} \\
\hline
\multirow{9}{*}{Lua Standard Libraries}
& base & 9 & 1 & 16 & 26 & 35\% \\
\cline{2-7}
& coroutine & 1 & 0 & 6 & 7 & 14\% \\
\cline{2-7}
& package & 5 & 0 & 3 & 8 & 62\% \\
\cline{2-7}
& string & 14 & 0 & 2 & 16 & 88\% \\
\cline{2-7}
& utf8 & 6 & 0 & 0 & 6 & 100\% \\
\cline{2-7}
& table & 1 & 5 & 1 & 7 & 14\% \\
\cline{2-7}
& math & 26 & 0 & 1 & 27 & 96\% \\
\cline{2-7}
& io & 15 & 0 & 6 & 21 & 71\% \\
\cline{2-7}
& os & 10 & 0 & 1 & 11 & 91\% \\
\hline
\end{tabular}
\end{center}
\caption{Evaluation results for Lua Standard Libraries 5.3}
\label{tab:evallsl53}
\end{table}

@@

\section{Typed Lua Compiler}

The Typed Lua compiler is the last case study that we evaluated, and the
Table \ref{tab:evaltlc} summarizes the evaluation results.

\begin{table}[!ht]
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|}
\hline
\textbf{Case study} & \textbf{Module} & \textbf{easy} & \textbf{poly} & \textbf{hard} & \textbf{Total} & \textbf{\%} \\
\hline
\multirow{8}{*}{Typed Lua Compiler}
& tlast & 46 & 0 & 1 & 47 & 98\% \\
\cline{2-7}
& tltype & 65 & 0 & 0 & 65 & 100\% \\
\cline{2-7}
& tlst & 26 & 0 & 0 & 26 & 100\% \\
\cline{2-7}
& tllexer & 2 & 0 & 9 & 11 & 18\% \\
\cline{2-7}
& tlparser & 1 & 0 & 0 & 1 & 100\% \\
\cline{2-7}
& tldparser & 1 & 0 & 0 & 1 & 100\% \\
\cline{2-7}
& tlchecker & 2 & 0 & 0 & 2 & 100\% \\
\cline{2-7}
& tlcode & 1 & 0 & 0 & 1 & 100\% \\
\hline
\end{tabular}
\end{center}
\caption{Evaluation results for Typed Lua Compiler}
\label{tab:evaltlc}
\end{table}

The \texttt{tlast} module implements the abstract syntax tree that
the compiler uses to perform the type checking.
We could not precisely type just one function because its return
type depend on the type of its input parameter.

The \texttt{tltype} module implements the types introduced by
Typed Lua and the subtyping and consistent-subtyping relationsr;
it was straightforward to type.

The \texttt{tlst} module implements the environment that the
compiler uses during the type checking, and it was also straightforward
to type.

The \texttt{tllexer} module defines common lexical rules for
the Typed Lua parser and the parser of the description files.
This module is hard to type because it exports LPeg \citep{lpeg,ierusalimschy2009lpeg}
patterns, which our type system cannot precisely type because it does
not support operator overloading.

The \texttt{tlparser} and \texttt{tldparser} implement the Typed Lua
parser and the parser of the description files, respectively.
Even though they use LPeg to implement the grammar rules,
they only export a parsing function that uses LPeg to parse a
string and return the corresponding AST.

The \texttt{tlchecker} and \texttt{tlcode} modules are also
straightforward to type.
The first traverses the AST to perform type checking,
while the latter traverses the AST to perform code generation.

Even though we could precisely type the interface of most of
the modules, we had some issues to write mutually recursive
functions that traverses the AST.
In this case we had to rely on the dynamic type to defer the
function declarations.

@@

\begin{itemize}
\item \citep{tidallock}
\item \citep{bonnaire-sergeant2012typed-clojure}
\item \citep{vitousek2014deg}
\item \citep{allende2013gts}
\item \citep{tobin-hochstadt2008ts} 
\item \citep{dart}
\item \citep{typescript}
\item \citep{bierman2014typescript}
\item \citep{politz2012semantics}
\end{itemize}
